### 2015-05-15 evaluation.tex
some fixes

\noindent\textbf{Accelerating \corral }
%
Table~\ref{tab:results} presents runtime results from using \whoop, \corral and \whoop + \corral to analyze our benchmarks, while Figure~\ref{fig:plot} plots the runtime speedups that \corral achieves using \whoop with the \yieldmr instrumentation. All reported times are in seconds and averages over three runs. \corral was configured with a time budget of 10 hours (T.O. denotes timeout) and a csb of 2, 5 and 9, respectively. Standalone \corral uses \yieldall, which instruments context-switches (i.e. \texttt{yield} statements) after all visible operations, while \whoop + \corral uses \yieldcoarse and \yieldmr, which instrument context-switches in a more fine-grained fashion (see \S\ref{corral}). The table also shows the number of context-switches per instrumentation (\#Y).

@@

Using the race-freedom guarantees from \whoop, we managed to significantly accelerate \corral in most of our benchmarks; the best results were achieved using \yieldmr. Figure~\ref{fig:plot} shows that most speedups using \yieldmr are between 1.5$\times$ and 10$\times$; while in ssu100, and pc8736x\_gpio respectively, we achieved a speedup of 12$\times$, and 20$\times$ respectively, using a csb of 9. We noticed that a higher csb typically results into greater speedups, when exploiting \whoop. This is expected, as a higher csb directly translates into a larger sequentialized program, and thus \whoop has more opportunities to reduce the sequentialization.
%
However, if \whoop fails to verify any race-related properties of a driver, it can slow down \corral. We noticed this in the sx8 driver, where \whoop does not verify any pairs, causing unnecessary overhead. Although \whoop verified 46 out of 47 memory regions of sx8, it still failed to speedup \corral using \yieldmr. \corral checks the only two pairs associated with sx8 quite quickly (21.4 seconds with csb of 9), so the overhead of running \whoop outweighs the benefits of fine-grained instrumentation.

\noindent\textbf{Other Tools }
%
We tried to compare \whoop with other similar approaches (see \S\ref{related}). However, we found this hard in practice: we downloaded Locksmith~\cite{pratikakis2006locksmith}, but could not get it to work with the 4.0 Linux drivers (the tool was last updated in 2007 and might miss some of the latest features); we also could not find source code or binaries for \cite{kahlon2007fast, kahlon2009semantic, das2015section}.

>>>>

\noindent\textbf{Accelerating \corral }
%
Table~\ref{tab:results} presents runtime results from using \whoop, \corral and \whoop + \corral to analyze our benchmarks, while Figure~\ref{fig:plot} plots the runtime speedups that \corral achieves using \whoop with the \yieldmr instrumentation. All reported times are in seconds and averaged over three runs. \corral was configured with a time budget of 10 hours (T.O. denotes timeout) and a csb of 2, 5 and 9. Standalone \corral uses \yieldall, which instruments context-switches (i.e. \texttt{yield} statements) after all visible operations, while \whoop + \corral uses \yieldcoarse and \yieldmr, which instrument context-switches in a more fine-grained fashion (see \S\ref{corral}). The table also shows the number of context-switches per instrumentation (\#Y).

@@

Using the race-freedom guarantees from \whoop, we managed to significantly accelerate \corral in most of our benchmarks; the best results were achieved using \yieldmr. Figure~\ref{fig:plot} shows that most speedups using \yieldmr are between 1.5$\times$ and 10$\times$; while in ssu100, and pc8736x\_gpio respectively, we achieved a speedup of 12$\times$, and 20$\times$ respectively, using a csb of 9. We noticed that a higher csb typically results into greater speedups when exploiting \whoop. This is expected, as a higher csb directly translates into a larger sequentialized program, and thus \whoop has more opportunities to reduce the sequentialization.
%
However, if \whoop fails to verify any race-related properties of a driver, it can slow down \corral. We noticed this in the sx8 driver, where \whoop does not verify any pairs, causing unnecessary overhead. Although \whoop verified 46 out of 47 memory regions of sx8, it still failed to speedup \corral using \yieldmr. \corral checks the only two pairs associated with sx8 quite quickly (21.4 seconds with csb of 9), so the overhead of running \whoop outweighs the benefits of fine-grained instrumentation.

\noindent\textbf{Other Tools }
%
We tried to compare \whoop with other similar approaches (see \S\ref{related}). However, we found this to be hard in practice: we downloaded Locksmith~\cite{pratikakis2006locksmith}, but could not get it to work with the 4.0 Linux drivers (the tool was last updated in 2007 and might be missing some of the latest features); we also could not find source code or binaries for \cite{kahlon2007fast, kahlon2009semantic, das2015section}.

### 2015-05-15 related.tex
Done it.
some fixes

Static race analysis is a promising alternative to dynamic techniques, which restrict analysis to the schedule chosen by a (possibly controlled) scheduler, providing limited coverage~\cite{musuvathi2008finding}. Warlock~\cite{sterling1993warlock} and LockLint~\cite{oracle2010locklint} are notable static race analyzers.  In comparison to \whoop, these tools rely heavily on user annotations.

Most related to our lockset analysis are the static lockset analyzers RELAY~\cite{voung2007relay} and Locksmith~\cite{pratikakis2006locksmith}. Both tools, though, have several limitations. RELAY found 5022 warnings when analyzing the Linux kernel, with only 25 of them being true data races. To tackle this issue, RELAY employs post-analysis unsound filters, but these can also filter out real bugs. Locksmith was successfully applied in several small Pthreads applications and 7 medium-sized Linux device drivers, but the authors reported that the tool was unable to run on several large programs, showcasing its limited scalability. \whoop aims to achieve scalability \emph{and} precision: the first, via novel symbolic pairwise lockset analysis; and the second, by accelerating \corral, an industrial-strength precise bug-finder.

Prior works on static race checking include: \cite{choi2002efficient}, which combines static analysis and runtime access caching to speedup dynamic race detection; \cite{kahlon2007fast}, which uses a divide-and-conquer algorithm that partitions all pointers of a program that do not alias in disjoint sets to achieve a scalable race analysis; \cite{kahlon2009semantic}, which uses abstract interpretation to achieve a sound partial-order reduction on the set of thread interleavings and statically reduce the number of false race warnings; and \cite{das2015section}, which employs inter-procedural alias analysis and verifiable user annotations to split programs into disjoint sections, based on non-communicating accesses of shared data, and eliminate redundant checks during dynamic race detection.

>>>>

Static race analysis is a promising alternative to dynamic techniques, which restrict analysis to the schedule chosen by a (possibly controlled) scheduler, providing limited coverage~\cite{musuvathi2008finding}. Warlock~\cite{sterling1993warlock} and LockLint~\cite{oracle2010locklint} are notable static race analyzers.  In comparison to \whoop, these tools rely heavily on user annotations.
%
Most related to our lockset analysis are the static lockset analyzers RELAY~\cite{voung2007relay} and Locksmith~\cite{pratikakis2006locksmith}. Both tools, though, have several limitations. RELAY found 5022 warnings when analyzing the Linux kernel, with only 25 of them being true data races. To tackle this issue, RELAY employs unsound post-analysis filters, but these can also filter out real bugs. Locksmith was successfully applied in several small Pthreads applications and 7 medium-sized Linux device drivers, but the authors reported that the tool was unable to run on several large programs, showcasing its limited scalability. \whoop aims to achieve scalability \emph{and} precision: the first, via novel symbolic pairwise lockset analysis; and the second, by accelerating \corral, an industrial-strength precise bug-finder.

Prior works on static race checking also include: \cite{choi2002efficient}, which combines static analysis and runtime access caching to speedup dynamic race detection; \cite{kahlon2007fast}, which uses a divide-and-conquer algorithm that partitions all pointers of a program that do not alias in disjoint sets to achieve a scalable race analysis; \cite{kahlon2009semantic}, which uses abstract interpretation to achieve a sound partial-order reduction on the set of thread interleavings and statically reduce the number of false race warnings; and \cite{das2015section}, which employs inter-procedural alias analysis and verifiable user annotations to split programs into disjoint sections, based on non-communicating accesses of shared data, and eliminate redundant checks during dynamic race detection.
